---
title: "2_centralTendancy"
author: "kithsiri Jayakody"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This mark down file is using to show the mean, median and the mode value of our dataset

```{r}
# dataset is loading and converting the character based data types to factors
df = read.csv('sperformance-dataset.csv',stringsAsFactors = F)

# converting the character type into factorial variables and see the conversion
indices <- c(1,2,4:6,9:14,18:23,34,38:43)

for(i in indices)
{
  df[[i]] <- as.factor(df[[i]])
}
```

calculating the mean, median and mode of few continous variable of the dataset - age column
```{r}
# age variable

mean(df$age)
median(df$age)

# there is no mode function with r. there fore create a user function to calculate mode of a data # set in R. This function takes the vector as input and gives the mode value as output.


# Create the function.
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

# calling the function and passing the age column to get the mode

result <- getmode(df$age)
print(result)
```

calculating the mean, median and mode of few continous variable of the dataset - mG1 ( maths first
period grade)
```{r}

# maths first grade  variable

mean(df$mG1)
median(df$mG1)


# calling the function and passing the age column to get the mode

result <- getmode(df$mG1)
print(result)
```
calculating the mean, median and mode of few continous variable of the dataset - pG1 ( protuguese first
period grade)

```{r}

# protuguese first grade  variable

mean(df$pG1)
median(df$pG1)


# calling the function and passing the age column to get the mode

result <- getmode(df$pG1)
print(result)

```



