---
title: "3_Variations"
author: "kithsiri Jayakody"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

loading the dataset and converting the character type variables to factors

```{r cars}
# dataset is loading and converting the character based data types to factors
setwd('D:/statistics')
df = read.csv('sperformance-dataset.csv',stringsAsFactors = F)

# converting the character type into factorial variables and see the conversion
indices <- c(1,2,4:6,9:14,18:23,34,38:43)

for(i in indices)
{
  df[[i]] <- as.factor(df[[i]])
}
```


calculating the variance, Standard deviation and quantiles with of few continous variable of the dataset - age variable

```{r}
var(df$age)
sd(df$age)
quantile(df$age)

```
calculating the variance, Standard deviation and quantiles with of few continous variable of the dataset - maths first grade variable

```{r}
var(df$mG1)
sd(df$mG1)
quantile(df$mG1)

```
calculating the variance, Standard deviation and quantiles with of few continuous variable of the dataset - protuguese first grade variable
```{r}
var(df$pG1)
sd(df$pG1)
quantile(df$pG1)

```

```