---
title: "6_Visuals_for_categorical_variables"
author: "kithsiri Jayakody"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(dplyr)
library(ggplot2)
```


this R markdown file consists of Visuals for categorical variables.

```{r}
# dataset is loading and converting the character based data types to factors
setwd('D:/statistics')
df = read.csv('sperformance-dataset.csv',stringsAsFactors = F)

# converting the character type into factorial variables and see the conversion
indices <- c(1,2,4:6,9:14,18:23,34,38:43)

for(i in indices)
{
  df[[i]] <- as.factor(df[[i]])
}
```

display the number of records based on sex. in here dplyr package was used to generate the bar plot

```{r}
# group the number of records based on sex 

df_sex = df %>% group_by(sex) %>% summarise(count=n())

# use the box plot to visualize the data

bp<- ggplot(df_sex, aes(x="", y=count, fill=sex))+
geom_bar(width = 1, stat = "identity")
bp


```
following code is used to generate the pie chart for the dataset with the sex category

```{r}


pie <- bp + coord_polar("y", start=0)
pie
```

following code block generate the pie chart based on internet usage or not category

```{r}

# group the number of records based on sex 

df_internet = df %>% group_by(internet) %>% summarise(count=n())

# use the box plot to visualize the data

bp<- ggplot(df_internet, aes(x="", y=count, fill=internet))+
geom_bar(width = 1, stat = "identity")
bp <- bp + coord_polar("y", start=0)
bp


```
following code display  the categorization based on father's job

```{r}

# group the number of records based on father's job 

df_Fjob = df %>% group_by(Fjob) %>% summarise(count=n())

# use the box plot to visualize the data

bp<- ggplot(df_Fjob, aes(x="", y=count, fill=Fjob))+
geom_bar(width = 1, stat = "identity")
bp <- bp + coord_polar("y", start=0)
bp


```



